---
title: "Autoregressive"
author: "Chris Simoes"
date: "4/7/2021"
output: html_document
---

### Script setup
Install packages, start the README.md, set the random seed  
```{r Script setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

ip <- installed.packages()
packages_used <- c("lattice", "tidyverse", "devtools", "here","viridisLite", "leaflet", "lme4")

ip <- installed.packages()
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}#END packages_used

here::i_am("analysis/Autoregressive.Rmd")

random_seed <- 835246
set.seed(random_seed)

```



## Spatial Correction 2D Autoregressive

Here I am modeling for spatial correction. The key is to calculate the R matrix which is a Kronecker product between correlation rows and correlation columns.<br>
Simulating trial with 4 varieties and 4 replications.

```{r spatial correction, echo = TRUE}
#Creating a matrix with RCBD trial
library(agricolae)
T=matrix("NA",16,4)
trt <- c("tr1","tr2","tr3","tr4")
rep <- 4
myRCBD <- design.rcbd(trt,r=rep, seed=-513, serie=1)
book2<- zigzag(myRCBD)

# Creating the trial matrix
j=1
for( i in 1:nrow(book2)){
  T[i,1] <- paste("",book2[i,3],sep="")
  T[i,3] <- book2[i,2]
  T[i,2] <- j
  j=j+1
  if(j>4){
    j=1
  }
}

T <-  noquote(T[order(T[,1]),]) # Sorting by treatment in this case genotypes

# Incidence Matrix
Xf=matrix(0,16,4)
Xf[,1]=c(1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0) # Variet 1 in each replication
Xf[,2]=c(0,0,0,0,1,1,1,1,0,0,0,0,0,0,0,0) # Variet 2 in each replication
Xf[,3]=c(0,0,0,0,0,0,0,0,1,1,1,1,0,0,0,0) # Variet 3 in each replication
Xf[,4]=c(0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1) # Variet 4 in each replication
Xf

# calculating the rank of the matrix
Matrix::rankMatrix(Xf) # Check that the rank is 4 and Xf has 4 columns. It works when getting the inverse matrix.

# Calculating X'X
XftXf = t(Xf)%*%(Xf)
XftXf

#Simulating data for each treatment
beta2=c(100,110,115,120) # this is a vector with 4 means for each variety
y=Xf%*%beta2+rnorm(16,0,2) # here I am simulating 16 values for y
Tnew <- noquote(data.frame(cbind(T[,1:3], y)))
colnames(Tnew) <- c("Treatment", "Col", "Row", "Yield")

source(here::here("code/prepareData.R"))

RCM = createCorrelation(Tnew$Row) #Calling correlation per row
CCM = createCorrelation(Tnew$Col) #Calling correlation per col

# Calculating variance
p=rep(1,nrow(Tnew)) #vector
s = as.numeric(Tnew$Yield)
varS = (t(s - mean(s))%*%(s - mean(s)))/(length(s)-1)

# Calculating the autoregressive vector
# varS * RCM %x% CCM (where %x% means kronecker product)
R = varS[1] * RCM %x% CCM 

corrected <- R %*% p #Corrected values per plot

finalTable <- cbind(Tnew, corrected)
knitr::kable(finalTable)

```


## Printig Heatmap Before Correction

```{r plotBefore, echo=TRUE}

z <- prepareHeatmap(finalTable$Yield)
coul <- viridis(100)
class(z) <- "numeric"
levelplot( z, xlab="Columns", ylab="Rows", main = "Before Correction", col.regions=coul)

```

## Printig Heatmap After Correction

```{r plotAfter, echo=TRUE}
w <- prepareHeatmap(finalTable$corrected)
coul <- viridis(100)
class(w) <- "numeric"
levelplot(w, xlab="Columns", ylab="Rows", main = "After Correction", col.regions=coul)

```


## Comparing Heritabilities

I am using lme4 in order to get genotypic and residual variances.

```{r heritability, echo=TRUE}

# Prepareing data for analysis
analysisData <- cbind(book2[order(book2$trt),],finalTable$Yield,finalTable$corrected)
colnames(analysisData) <- c("plot","block","trt","Yield","corrected")
analysisData$Yield <- as.numeric(analysisData$Yield)

#Running lme4 with treatment as random effect and block as fixed.
library("lme4") 
# First with no correction
model_yield <- lmer( Yield ~ (1|trt) + block, na.action = na.exclude, data=analysisData)
variance_yield = as.data.frame(VarCorr(model_yield))

gvar_yield <- variance_yield[1,"vcov"] # Genotypic variance
resvar_yield <- variance_yield[2,"vcov"] # residual variance

#Heritability before correction
h2_yield <- gvar_yield/(gvar_yield+resvar_yield)

# Here with corrected values
model_corrected <- lmer(corrected ~ (1|trt) + block, na.action = na.exclude, data=analysisData)
variance_corrected = as.data.frame(VarCorr(model_corrected))
gvar_corrected <- variance_corrected[1,"vcov"] # Genotypic variance
resvar_corrected <- variance_corrected[2,"vcov"] # Residual variance

# Heritability after correction
h2_corrected <- gvar_corrected/(gvar_corrected+resvar_corrected)

#Results
h2_yield
h2_corrected



```


